---
title: "R Notebook"
output: html_notebook
---


```{r}
library(survival)
library(broom)
setwd('E:/Data/masterarbeit/Data/Tables/rhem')
```
# num_referrals + exact_repetition + unordered_repetition + partial_repetition_1 + partial_repetition_2 + partial_repetition_3 + partial_repetition_4 + partial_repetition_5 + partial_repetition_6 + partial_repetition_7 + partial_repetition_8 + partial_repetition_9 + partial_repetition_10 + unordered_partial_repetition_0 + unordered_partial_repetition_1 + unordered_partial_repetition_2 + unordered_partial_repetition_3 + unordered_partial_repetition_4 + unordered_partial_repetition_5 + unordered_partial_repetition_6 + unordered_partial_repetition_7 + unordered_partial_repetition_8 + unordered_partial_repetition_9 + unordered_partial_repetition_10 + exact_reciprocation + referee_indegree + referral_outdegree + transitive_tie + cyclical_tie + triangle_effect_children + triangle_effect_parents

```{r}
nomination.events <- read.csv('bucharest_eventlist_nomination.csv', sep=',')
nomination.events[,c(which(names(nomination.events) == 'num_referrals'):which(names(nomination.events) == 'triangle_effect_parents'))] <- log1p(nomination.events[,c(which(names(nomination.events) == 'num_referrals'):which(names(nomination.events) == 'triangle_effect_parents'))])
nomination.events[,c(which(names(nomination.events) == 'num_referrals'):which(names(nomination.events) == 'triangle_effect_parents'))] <- scale(nomination.events[,c(which(names(nomination.events) == 'num_referrals'):which(names(nomination.events) == 'triangle_effect_parents'))])
nomination.events.surv <- Surv(time=rep(1, nrow(nomination.events)), event=nomination.events$IS_OBSERVED)
nomination.events.rem <- coxph(nomination.events.surv ~ num_referrals + exact_repetition + unordered_partial_repetition_1 + partial_repetition_0 + partial_repetition_1 + exact_reciprocation + referee_indegree + referral_outdegree + cyclical_tie + triangle_effect_children + referee_age + referrals_avg_age + referee_referrals_age_difference + referrals_age_difference + referee_gender + referrals_avg_gender + referee_referrals_gender_difference + referrals_gender_difference + referee_medical + referrals_avg_medical + referee_referrals_medical_difference + referrals_medical_difference + referee_isco08_label + referrals_avg_isco08_label + referee_referrals_isco08_label_difference + referrals_isco08_label_difference + strata(EVENT_INTERVAL), data=nomination.events)
# unordered repetition, partial repetition 3-10, unordered partial repetition 2-10, triangle_effect_parents NAN
```

```{r}
options(width=300)
summary(nomination.events.rem)
```
```{r}
write.csv(tidy(nomination.events.rem), file='../../../Thesis/tables/rhem/bucharest_rhem.csv')
```


```{r}
nomination.events.cond.referee <- read.csv('bucharest_eventlist_nomination_cond_referee.csv', sep=',')
nomination.events.cond.referee[,c(which(names(nomination.events.cond.referee) == 'num_referrals'):which(names(nomination.events.cond.referee) == 'triangle_effect_parents'))] <- log1p(nomination.events.cond.referee[,c(which(names(nomination.events.cond.referee) == 'num_referrals'):which(names(nomination.events.cond.referee) == 'triangle_effect_parents'))])
nomination.events.cond.referee[,c(which(names(nomination.events.cond.referee) == 'num_referrals'):which(names(nomination.events.cond.referee) == 'triangle_effect_parents'))] <- scale(nomination.events.cond.referee[,c(which(names(nomination.events.cond.referee) == 'num_referrals'):which(names(nomination.events.cond.referee) == 'triangle_effect_parents'))])
nomination.events.cond.referee.surv <- Surv(time=rep(1, nrow(nomination.events.cond.referee)), event=nomination.events.cond.referee$IS_OBSERVED)
nomination.events.cond.referee.rem <- coxph(nomination.events.cond.referee.surv ~ num_referrals + unordered_partial_repetition_1 + unordered_partial_repetition_2 + exact_reciprocation + referee_indegree + referral_outdegree + cyclical_tie + triangle_effect_children + referee_age + referrals_avg_age + referee_referrals_age_difference + referrals_age_difference + referee_gender + referrals_avg_gender + referee_referrals_gender_difference + referrals_gender_difference + referee_medical + referrals_avg_medical + referee_referrals_medical_difference + referrals_medical_difference + referee_isco08_label + referrals_avg_isco08_label + referee_referrals_isco08_label_difference + referrals_isco08_label_difference + strata(EVENT_INTERVAL), data=nomination.events.cond.referee)
# exact repetition, unordered repetition, partial repetition 2-10, unordered partial repetition 0-10, transitive_tie, triangle_effect_parents NAN
```

```{r}
options(width=300)
summary(nomination.events.cond.referee.rem)
```
```{r}
write.csv(tidy(nomination.events.cond.referee.rem), file='../../../Thesis/tables/rhem/bucharest_rhem_cond_sender.csv')
```

```{r}
nomination.events.cond.referral <- read.csv('bucharest_eventlist_nomination_cond_referral.csv', sep=',')
nomination.events.cond.referral[,c(which(names(nomination.events.cond.referral) == 'num_referrals'):which(names(nomination.events.cond.referral) == 'triangle_effect_parents'))] <- log1p(nomination.events.cond.referral[,c(which(names(nomination.events.cond.referral) == 'num_referrals'):which(names(nomination.events.cond.referral) == 'triangle_effect_parents'))])
nomination.events.cond.referral[,c(which(names(nomination.events.cond.referral) == 'num_referrals'):which(names(nomination.events.cond.referral) == 'triangle_effect_parents'))] <- scale(nomination.events.cond.referral[,c(which(names(nomination.events.cond.referral) == 'num_referrals'):which(names(nomination.events.cond.referral) == 'triangle_effect_parents'))])
nomination.events.cond.referral.surv <- Surv(time=rep(1, nrow(nomination.events.cond.referral)), event=nomination.events.cond.referral$IS_OBSERVED)
nomination.events.cond.referral.rem <- coxph(nomination.events.cond.referral.surv ~ num_referrals + unordered_partial_repetition_1 + partial_repetition_0 + exact_reciprocation + referee_indegree + referral_outdegree + cyclical_tie + triangle_effect_children + referee_age + referrals_avg_age + referee_referrals_age_difference + referrals_age_difference + referee_gender + referrals_avg_gender + referee_referrals_gender_difference + referrals_gender_difference + referee_medical + referrals_avg_medical + referee_referrals_medical_difference + referrals_medical_difference + referee_isco08_label + referrals_avg_isco08_label + referee_referrals_isco08_label_difference + referrals_isco08_label_difference + strata(EVENT_INTERVAL), data=nomination.events.cond.referral)
# exact_repetition, unordered repetition, partial repetition 2-10, unordered partial repetition 1-10, transitive_tie, triangle_effect_parents NAN
```

```{r}
options(width=300)
summary(nomination.events.cond.referral.rem)
```
```{r}
write.csv(tidy(nomination.events.cond.referral.rem), file='../../../Thesis/tables/rhem/bucharest_rhem_cond_receiver.csv')
```